---
title: "R Notebook"
output: html_notebook
---


## following tutorial
https://github.com/StatsGary/NHSRComm_TidyModels_Recipes
https://hutsons-hacks.info/tidymodels-hands-on-for-the-nhs-r-community-building-a-supervised-ml-classification-model-from-scratch

file:///F:/learning/NHSRComm_TidyModels_Recipes-main/Building_a_TidyModel_for_classification_from_scratch_full_tutorial.html#1_Starting_out_with_TidyModels 

## Packages

```{r}
#load packages ----
library(tidyverse)
library(NHSRdatasets)
library(tidymodels)
library(themis)
library(plotly)
```


## Step one - Importing data into the R environment

```{r}
# Read in the data ----
strand_pat <- NHSRdatasets::stranded_data %>% 
  setNames(c("stranded_class", "age", "care_home_ref_flag", "medically_safe_flag", 
             "hcop_flag", "needs_mental_health_support_flag", "previous_care_in_last_12_month", "admit_date", "frail_descrip")) %>% 
  mutate(stranded_class = factor(stranded_class)) %>% 
  drop_na()

print(head(strand_pat))
```


## Step Two - Analysing the Class Imbalance

```{r}
class_bal_table <- table(strand_pat$stranded_class)
prop_tab <- prop.table(class_bal_table)
upsample_ratio <- (class_bal_table[2] / sum(class_bal_table))

prop_tab
upsample_ratio
```

## Step Three - Observe data structures

```{r}
strand_pat$admit_date <- as.Date(strand_pat$admit_date, format="%d/%m/%Y") #Format date to be date to work with recipes steps
factors <- names(select_if(strand_pat, is.factor))
numbers <- names(select_if(strand_pat, is.numeric))
characters <- names(select_if(strand_pat, is.character))

print("factors"); factors 
print("numbers"); numbers 
print("characters"); characters
```

## Step Four - Using Rsample to create ML data partitions

```{r}
# Partition into training and hold out test / validation sample
set.seed(123)
split <- rsample::initial_split(strand_pat, prop=3/4)
train_data <- rsample::training(split)
test_data <- rsample::testing(split)
```

## Step Five - Creating your first Tidy Recipe

```{r}
stranded_rec <- 
  recipe(stranded_class ~ ., data=train_data) %>% 
  # The stranded class is what we are trying to predict and we are using the training data
  
  step_date(admit_date, features = c("dow", "month")) %>% 
  #Recipes step_date allows for additional features to be created from the date 
  step_rm(admit_date) %>% 
  #Remove the date, as we have created features off of it, if left in the dreaded multicolinearity may be present
  
  themis::step_upsample(stranded_class, over_ratio = as.numeric(upsample_ratio)) %>%  
  #SMOTE recipe step to upsample the minority class i.e. stranded patients
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  #Automatically created dummy variables for all categorical variables (nominal)
  step_zv(all_predictors()) %>% 
  #Get rid of features that have zero variance
  step_normalize(all_predictors()) #ML models train better when the data is centered and scaled

print(stranded_rec) #Terminology is to use recipe
```

## Step Six - Get Parsnipping

### Instantiate the model

```{r}
lr_mod <- 
  parsnip::logistic_reg() %>% 
  set_engine("glm")

print(lr_mod)
```

### Creating the model workflow
```{r}
# Create model workflow
strand_wf <- 
  workflow() %>% 
  add_model(lr_mod) %>% 
  add_recipe(stranded_rec)

print(strand_wf)
```

### Fitting the workflow to our data
```{r}
# Create the model fit
strand_fit <- 
  strand_wf %>% 
  fit(data = train_data)
```


### Extracting the fitted data
```{r}
strand_fitted <- strand_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

print(strand_fitted)
```

###  Create custom plot to visualise significance utilising p values
```{r}
# Add significance column to tibble using mutate
strand_fitted <- strand_fitted  %>% 
  mutate(Significance = ifelse(p.value < 0.05, "Significant", "Insignificant")) %>% 
  arrange(desc(p.value)) 

#Create a ggplot object to visualise significance
plot <- strand_fitted %>% 
  ggplot(data = strand_fitted, mapping = aes(x=term, y=p.value, fill=Significance)) +
  geom_col() + theme(axis.text.x = element_text(
                                        face="bold", color="#0070BA", 
                                        size=8, angle=90)
                                                ) + labs(y="P value", x="Terms", 
                                                         title="P value significance chart",
                                                         subtitle="A chart to represent the significant variables in the model",
                                                         caption="Produced by Gary Hutson")

#print("Creating plot of P values")
#print(plot)
plotly::ggplotly(plot)
```














